---
title: "Reproducible Research Project 1"
output: github_document
---

## Introducing the Project
In this project, we will be analyzing an anonymous individual's walking routine over the following time interval: **October 1st 2012 - November 30th 2012**. Specifically, we will assess the number of steps taken over five minute interval measurements. 

##Reading the CSV file

The "activity.csv" file that was provided by Coursera is already relatively easy to work with. I created the following function to read the file into my workspace:

```{r}
#set appropriate working directory (location depends on individual)
setwd("~/R Material/Coursera/Course 5 Reproducible Research/Week 1/repdata_data_activity")

#Read in csv file, cleaning up column names and coercing 2nd column into a Date class
ReadFile <- function(){
      return(read.csv("activity.csv", colClasses = c("integer","Date", "integer"),
                      col.names = c("Steps", "Date", "Interval")))
}

ActivityFile <- ReadFile()
head(ActivityFile)
```

Before moving onto the assignment questions, we will look at a summary of the data table
```{r}
summary(ActivityFile)
```

There are 2304 NA values for the Steps column, so we will keep this in mind when we move forward with the analysis. 

##Finding the mean total number of steps per day


